---
categories:  
- ""    #the front matter should be like the one found in, e.g., blog2.md. It cannot be like the normal Rmd we used
- ""
date: "2022-09-14"
description: A graph depicting the best and worst countries regarding renewable energy production # the title that will show up once someone gets to this page
draft: false
image: / #save picture in \static\img\blogs. Acceptable formats= jpg, jpeg, or png . Your iPhone pics wont work
keywords: ""
slug: brexit # slug is the shorthand URL address... no spaces plz
title: "Share of renewable energy production: tops and flops"
---


```{r load-libraries,include=FALSE, cache = FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(GGally)
library(readxl)
library(here)
library(skimr)
library(janitor)
library(broom)
library(tidyquant)
library(infer)
library(openintro)
library(countrycode)
```



The National Bureau of Economic Research (NBER) has a a very interesting
dataset on the adoption of about 200 technologies in more than 150
countries since 1800. This is the[Cross-country Historical Adoption of
Technology (CHAT) dataset](https://www.nber.org/research/data/cross-country-historical-adoption-technology).

The following is a description of the variables

| **variable** | **class** | **description**                |
|--------------|-----------|--------------------------------|
| variable     | character | Variable name                  |
| label        | character | Label for variable             |
| iso3c        | character | Country code                   |
| year         | double    | Year                           |
| group        | character | Group (consumption/production) |
| category     | character | Category                       |
| value        | double    | Value (related to label)       |

```{r,load_technology_data}

technology <- readr::read_csv(here::here('data/technology.csv'))

#get all technologies
labels <- technology %>% 
  distinct(variable, label)

# Get country names using 'countrycode' package
technology <- technology %>% 
  filter(iso3c != "XCD") %>% 
  mutate(iso3c = recode(iso3c, "ROM" = "ROU"),
         country = countrycode(iso3c, origin = "iso3c", destination = "country.name"),
         country = case_when(
           iso3c == "ANT" ~ "Netherlands Antilles",
           iso3c == "CSK" ~ "Czechoslovakia",
           iso3c == "XKX" ~ "Kosovo",
           TRUE           ~ country))

#make smaller dataframe on energy
energy <- technology %>% 
  filter(category == "Energy")

# download CO2 per capita from World Bank using {wbstats} package
# https://data.worldbank.org/indicator/EN.ATM.CO2E.PC
# co2_percap <- wb_data(country = "countries_only", 
#                       indicator = "EN.ATM.CO2E.PC", 
#                       start_date = 1970, 
#                       end_date = 2022,
#                       return_wide=FALSE) %>% 
#   filter(!is.na(value)) %>% 
#   #drop unwanted variables
#   select(-c(unit, obs_status, footnote, last_updated))

co2_percap <- read_csv(here::here("data/co2percap.csv"),skip = 4) %>% 
  janitor::clean_names()


```

This is a very rich data set, not just for energy and CO2 data, but for many other technologies. In our case, we just need to produce a couple of graphs-- at this stage, the emphasis is on data manipulation, rather than making the graphs gorgeous.

First, produce a graph with the countries with the highest and lowest % contribution of renewables in energy production. This is made up of `elec_hydro`, `elec_solar`, `elec_wind`, and `elec_renew_other`. You may want to use the *patchwork* package to assemble the two charts next to each other.
 
```{r min-max_renewables, echo=FALSE, out.width="100%"}
library(patchwork)

x <-  energy %>%
  pivot_wider(names_from=variable, values_from=value) %>%
  filter(year==2019) %>% 
  group_by(country) %>%
  summarise(elec_hydro=sum(elec_hydro, na.rm=TRUE),
           elec_solar=sum(elec_solar, na.rm=TRUE),
           elec_wind= sum(elec_wind, na.rm=TRUE),
           elec_renew=sum(elec_renew_other, na.rm=TRUE),
           elecprod = elecprod)%>%
  ungroup() %>% 
  mutate(percentage = (elec_hydro + elec_solar + elec_wind + elec_renew) / elecprod, 
         percentage = if_else(is.na(percentage),0,percentage), 
         percentage = if_else(is.infinite(percentage),0,percentage)) %>% 
  filter(elecprod > 0)
  
top_20 <- x%>%  
  mutate(country = fct_reorder(country,percentage)) %>% 
  slice_max(percentage, n = 20) %>% 
  ggplot(aes(percentage,country))  + 
  geom_col()+ 
  theme_bw()



bottom_20 <- x%>%  
  mutate(country = fct_reorder(country,percentage)) %>%
  slice_min(percentage, n = 20) %>% 
  ggplot(aes(percentage,country))  + 
  geom_col()+ 
  theme_bw()


top_20 + bottom_20 
  

```
